---
title: "641 Project"
subtitle: "Fitting time series models to oil prices"
author: "Chris Flanagan, Suraj Nair, Gregory Yampolsky, Yehao Feng, Elliot Frost"
output: html_notebook
---


### Data cleaning and preparation

```{r}
data = read.csv('/Users/chrisflanagan/Desktop/641/project/DCOILWTICO.csv')
head(data)
```

rename columns to date and oil price (per barrel)

```{r}
colnames(data) = c('date','price')
head(data)
```

We need to change the type of the price column from chr to int
```{r}
# cast from chr to int 
data$price = as.integer(data$price)
typeof(data$price)
```

We need to check for missing values
```{r}
#C heck for missing values
sum(is.na(data$price))
```

We have 327 missing values, We will replace them with the mean of the column
```{r}
# Replace missing values with mean
data$price[is.na(data$price)] = mean(data$price, na.rm = TRUE) 
sum(is.na(data$price))
```

### Visualization
```{r}
# Visualize time series
plot(data$price, type = 'l')
```


We will check for normality
```{r}
# Check for normality
hist(data$price)
```

The data does not appear to be normal

## Check for staionarity


we will check for stationarity with an ACF plot
```{r}
acf(data$price)
pacf(data$price)
```

The ACF plot implies the data is not stationary.

We will also do an ADF test to check for stationarity
```{r}
library(tseries)
adf.test(data$price)
```

Since our p-value is > .05, we reject our alternative hypothesis. Our dataset is not stationary.

We will perform stationarity through differenceing
```{r}
# perform transformation
dseries = diff(data$price)
```

## Finding models

```{r}
acf(data$price,lag.max = 20)
acf(dseries, lag.max = 20)
pacf(data$price, lag.max = 20)
pacf(dseries, lag.max = 20)
```

Our ACF and PACF plots seem to imply an MA(1) model because the PACF plot tails off and the ACF plot cuts off at lag 1

We will fit an MA(1) model

im not sure to fit the differenced data or the original data
```{r}
ma1 = arima(dseries,order=c(0,0,1))
plot(ma1)
```

```{r}
arima(ma1,order=c(0,0,1),method='CSS')
arima(ma1,order=c(0,0,1),method='ML')
```
```{r}
plot(ma1,type = 'l')
```



